Meeting the Challenges of Modeling Astrophysical Thermonuclear Explosions: \castro, \maestro, and the \amrex\ Astrophysics Suite

### 2017-10-30 astronum_2017/astronum-amrex.tex
Small changes after one pass.

Among the most significant open questions for SNe Ia is the identity of the
progenitor.  About 20 years ago, the community had
mostly converged upon the single-degenerate scenario---a
Chandrasekhar-mass C/O white dwarf that accretes from its companion,
eventually leading to runaway at the center that burns through the
star (see~\cite{hillebrandtniemeyer2000} for the state of the field at
that time).  Since then, a wealth of observations has indicated that
there is a lot of diversity in SNe Ia, and searches for progenitor
systems have strongly suggested that Chandra-mass white dwarfs cannot
explain most SNe Ia.  Today, merging white dwarfs (the double
degenerate scenario) have perhaps become the most popular model.
Other progenitors, like He burning on the surface of a sub-Chandra
white dwarf, have also seen interest in explaining some of the observed
diversity.  See~\cite{araa-maoz} for a review.

There are open questions in all of these scenarios that can be
addressed through simulation.  For the Chandra and sub-Chandra models,
what is the distribution (spatial and temporal) of the hotspots set up
by turbulent convection that give rise to burning fronts?  A
longstanding question with the Chandra model is whether a deflagration
can transition to a detonation during the burning front propagation
through the star.  For the sub-Chandra double detonation model, is it
not clear whether it is possible to create a detonation in the thin
surface He layer.  For double degenerates, it is still unresolved
whether the burning takes place promptly or after some delay.  However
it proceeds, we need to avoid an accretion-induced collapse to a
neutron star.  For many of these investigations, we need to address
numerical issues such as whether it is possible to accurately model
the ignition of a detonation with the spatial resolution we can
attain.  These are some of the questions we seek to answer.

@@

\maestro~\cite{MAESTRO:Multilevel} is a low Mach number stellar
hydrodynamics code designed for efficiently modeling convection in
stars.  \maestro\ decomposes the state variables into a one-dimensional
hydrostatic base state and a three-dimensional Cartesian state that
models the deviation from hydrostatic equilibrium.
A constraint equation is derived by
requiring that the pressure everywhere be close to the background
hydrostatic pressure.  The constraint acts to enforce instantaneous
acoustic equilibration, effectively filtering soundwaves from the
system, while retaining the compressibility effects due to the
background stratification of the star and local heat release, as well
as the hydrostatic adjustment of the star.  In this fashion, it is
more general than traditional anelastic methods.  The timestep
constraint for these equations depends only on the fluid velocity, not
the sound speed, enabling much larger timesteps than compressible
codes for highly subsonic flows.

>>>>

Among the most significant open questions for SNe Ia is the identity of the
progenitor.  About 20 years ago, the community had
mostly converged upon the single-degenerate scenario---a
Chandrasekhar-mass C/O white dwarf that accretes from its companion,
eventually leading to runaway at the center that burns through the
star (see~\cite{hillebrandtniemeyer2000} for the state of the field at
that time).  Since then, a wealth of observations has indicated that
there is a lot of diversity in SNe Ia, and searches for progenitor
systems have strongly suggested that Chandra-mass white dwarfs cannot
explain most SNe Ia.  Today, merging white dwarfs (the double
degenerate scenario) has perhaps become the most popular model.
Other progenitors, like He burning on the surface of a sub-Chandra
white dwarf, have also seen interest in explaining some of the observed
diversity.  See~\cite{araa-maoz} for a review.

There are open questions in all of these scenarios that can be
addressed through simulation.  For the Chandra and sub-Chandra models,
what is the distribution (spatial and temporal) of the hotspots set up
by turbulent convection that give rise to burning fronts?  A
longstanding question with the Chandra model is whether a deflagration
can transition to a detonation during the burning front propagation
through the star.  For the sub-Chandra double detonation model, it is
not clear whether it is possible to create a detonation in the thin
surface He layer.  For double degenerates, it is still unresolved
whether the burning takes place promptly or after some delay.  However
it proceeds, we need to avoid an accretion-induced collapse to a
neutron star.  For many of these investigations, we need to address
numerical issues such as whether it is possible to accurately model
the ignition of a detonation with the spatial resolution we can
attain.  These are some of the questions we seek to answer.

@@

\maestro~\cite{MAESTRO:Multilevel} is a low Mach number stellar
hydrodynamics code designed for efficiently modeling convection in
stars.  \maestro\ decomposes the state variables into a one-dimensional
hydrostatic base state and a three-dimensional Cartesian state that
models the deviation from hydrostatic equilibrium.
A constraint equation is derived by
requiring that the pressure everywhere is close to the background
hydrostatic pressure.  The constraint acts to enforce instantaneous
acoustic equilibration, effectively filtering soundwaves from the
system, while retaining the compressibility effects due to the
background stratification of the star and local heat release, as well
as the hydrostatic adjustment of the star.  In this fashion, it is
more general than traditional anelastic methods.  The timestep
constraint for these equations depends only on the fluid velocity, not
the sound speed, enabling much larger timesteps than compressible
codes for highly subsonic flows.
